Category-based and location-based volitional covert attention affect memory at different timescales

### 2023-03-25 paper/main.tex
working on results section

\newcommand{\familiarity}{S1}
\newcommand{\fixation}{S2}
\newcommand{\timecourseCuedCategory}{S3}
\newcommand{\timecourseUncuedCategory}{S4}
\newcommand{\timecourseCuedLocation}{S5}
\newcommand{\timecourseUncuedLocation}{S6}
\newcommand{\timecourseNovel}{S7}
\newcommand{\timecourseDiff}{S8}

@@

@@

@@

We ran two volitional covert attention experiments; in the Sustained Attention
Experiment we asked participants to \textit{sustain} the focus of their
attention over a succession of 10 stimulus presentations per block whereas in
the Variable Attention Experiment we asked participants to \textit{vary} their
focus of attention with each new stimulus (also for a total of 10 stimulus
presentations per block). Each stimulus comprised a pair of composite images
(one on the left and one on the right side of the display), where each
composite comprised an equal blend of a unique face and a unique place image.
We followed the presentation phases of each experimental block with a memory
phase, where participants performed a recognition memory task by rating the
familiarity of previously experienced and novel face and place images (see
\textit{Materials and methods}, Fig.~\ref{fig:paradigm}).

We considered each image we probed in the memory phase with respect to whether
(and how) the participant had experienced and/or attended to that image during
the presentation phase (Fig.~\ref{fig:familiarity}). For example, if a given
face image had been presented as part of the composite image on the right of
the screen following an \texttt{attend to the face on the right} instruction,
then that face image would have been the focus of the participant's attention
during one stimulus presentation. The face image on the left (unattended side)
during that trial would have been captured by feature-based attention, but not
by location-based attention; and the place image (unattended category) on the
right during that trial would have been captured by location-based attention,
but not by feature-based attention. The place image on the left during that
trial remained outside of the focus of both feature-based and location-based
attention. In this way, we categorized each of the images that participants
experienced during the presentation phases of each experiment in terms of
whether they fell under the scope of feature-based and/or location-based
attention. We also categorized novel face and place images that we asked
participants to judge during the memory phase (i.e., images that the
participant hadn't seen before) as belonging to the attention-cued or uncued
category. These novel images were intended to serve as a baseline for
comparison. For example, we sought to measure how people rated the familiarity
of new images in general. We also used differences in familiarity ratings
between novel images of the attention-cued versus uncued categories to evaluate
potential \textit{prospective} affects of modulating the focus of feature-based
attention.

\begin{figure*}[tp]
  \centering
  \includegraphics[width=0.75\textwidth]{figures/familiarity}

  \caption{\textbf{Familiarity ratings for attended, unattended, and novel
  stimuli.} Each split violin plot displays the distribution of
  within-participant average familiarity ratings given to faces (left, darker
  colors) and scenes (right, lighter colors) during the memory phases of each
  experiment. As shown in the legend (bottom), the colors indicate whether each
  image had been viewed during the presentation phase (\texttt{old}) or not;
  whether the given images matched the \texttt{attended category}; and/or
  whether the given images matched the \texttt{attended location}. The colored
  lines above each set of violin plots denote statistical differences
  (\textbf{pos}itive or \textbf{neg}ative differences in mean, collapsing over
  image category, assessed via two-tailed $t$-tests) between the distributions
  centered on the endpoints of each line. The line thicknesses denote
  $p$-values as indicated in the legend. Asterisks denote differences between
  the face versus scene distributions (assessed via two-tailed $t$-tests).
  Panel \textbf{a.} displays results from the Sustained Attention Experiment
  and panel \textbf{b.} displays results from the Variable Attention
  Experiment. Figure~\familiarity~displays these results broken down by
  participant cohort.}

\label{fig:familiarity}
\end{figure*}

Participants in both experiments rated stimuli they attended as more familiar
than unattended or novel stimuli (Fig.~\ref{fig:familiarity}; Sustained
Attention Experiment: $t(59) = 13.42, p = 1.39 \times 10^{-19}, d = 1.73$;
Variable Attention Experiment: $t(52) = 12.40, p = 3.70 \times 10^{-17}, d =
1.70$). Participants in both experiments also rated attended scene stimuli as
more familiar than attended face stimuli (Sustained Attention Experiment:
$t(59) = 7.69, p = 1.85 \times 10^{-10}, d = 0.99$; Variable Attention
Experiment: $t(52) = 7.08, p = 3.63 \times 10^{-9}, d = 0.97$). These findings
indicate that aspects of experience that are captured by the focus of attention
are easier to recognize later, and that attention may preferentially benefit
memory for some aspects of experience to a greater extent than other aspects of
experience.

Participants in the Sustained Attention Experiment experiment also rated
unattended images that matched the attended category as more familiar than
unattended images that did not match the attended category (\textit{incidental
feature-based attention}; $t(59) = 7.17, p = 1.40 \times 10^{-9}, d = 0.93$).
This pattern did not hold for participants in the Variable Attention Experiment
($t(52) = 0.44, p = 0.66, d = 0.06$). Participants in both experiments
displayed an incidental effect of location-based attention, whereby they rated
unattended-category images at the attended location as more familiar than
unattended-category images at the unattended location (Sustained Attention
Experiment: $t(59) = 4.32, p = 6.05 \times 10^{-5}, d = 0.56$; Variable
Attention Experiment: $t(52) = 2.57, p = 0.01, d = 0.35$). This indicates that
aspects of our experience that incidentally overlap with the focus of our
location-based attention are easier to recognize later, even if those aspects
do not fall within the intentional focus of our feature-based attention.

We defined the \textit{memory benefit of feature-based attention} as the
difference in mean familiarity ratings of attended stimuli and unattended
stimuli from the attended category versus the mean familiarity ratings of
unattended stimuli from the unattended category. Similarly, we defined the
\textit{memory benefit of location-based attention} as the difference in mean
familiarity ratings of attended stimuli and unattended stimuli from the
attended location versus the mean familiarity ratings of unattended stimuli
from the unattended location. Participants in the Sustained Attention
Experiment displayed a larger memory benefit of feature-based attention than
location-based attention ($t(59) = 3.25, p = 1.91 \times 10^{-3}, d = 0.42$),
whereas participants in the Variable Attention Experiment displayed a larger
memory benefit of location-based attention ($t(52) = 1.70, p = 9.99 \times
10^{-2}, d = 0.23$). This appeared to be driven in part by a response bias
towards attended-category images. For example, relative to the most recent
attention cue, participants in the Sustained Attention Experiment rated
attended-category novel images as more familiar than unattended-category novel
images ($t(59) = 7.04, p = 2.35 \times 10^{-9}, d = 0.91$), whereas
participants in the Variable Attention Experiment responded similarly to
attended-category and unattended-category novel images ($t(52) = 0.23, p =
0.82, d = 0.03$). The size of this attended-category response bias for novel
images was reliably smaller than the memory benefits of feature-based attention
in both experiments (Sustained Attention Experiment: $t(59) = 7.59, p = 2.68
\times 10^{-10}, d = 0.98$; Variable Attention Experiment: $t(52) = 6.74, p =
1.31 \times 10^{-8}, d = 0.93$). This indicates that tuning the focus of
feature-based attention with respect to how ongoing experience is encoded in
memory takes longer than the duration of a single stimulus presentation
(including fixation, roughly 5.5~s). Further, once feature-based attention
comes ``online,'' it affects how \textit{new} stimuli are processed.

We also compared the relative sizes of the memory benefits of feature-based
attention and location-based attention across the two experiments. Participants
in the Sustained Attention Experiment displayed a larger difference between
these two effects than did participants in the Variable Attention Experiment
($t(111) = 3.48, p = 7.24 \times 10^{-4}, d = 0.66$). Further, the difference
in mean familiarity ratings of attended-location images versus unattended
images was greater for participants in the Variable Attention Experiment
($t(111) = 2.32, p = 0.02, d = 0.44$). The memory benefit of location-based
attention reliably distinguished between attended-category versus unattended
category images for participants in both experiments (Sustained Attention
Experiment: $t(59) = 9.50, p = 1.68 \times 10^{-13}, d = 1.23$; Variable
Attention Experiment: $t(52) = 9.72, p = 2.76 \times 10^{-13}, d = 1.33$). This
indicates that tuning the focus of location-based attention with respect to how
ongoing experience is encoded into memory occurs faster than the duration of a
single stimulus presentation.

We also compared reaction times to probes from the attended versus unattended
side. Participants in both experiments responded slightly faster on average to
attended-side probes, although the differences did not reach significance
(Sustained Attention Experiment: $t(59) = 1.03, p = 0.31, d = 0.13$; Variable
Attention Experiment: $t(52) = 0.65, p = 0.52, d = 0.09$). Participants in the
Sustained Attention Experiment displayed a slightly larger (and not
statistically significant) difference in attended-side versus unattended-side
reaction times than participants in the Variable Attention Experiment ($t(111)
= .87, p = 0.38, d = 0.20$). Because none of these reaction time effects were
statistically reliable, we are unable to draw meaningful conclusions about
reaction times to the attention probes on the attended versus unattended side,
nor are we able to draw meaningful conclusions about differences in those
reaction times that might be due to sustained versus variable attention.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figures/exps1_2_xgaze.pdf}

  \caption{\textbf{Horizontal coordinates of visual fixations.} Each violin
  plot displays a distribution of participant-wise average horizontal gaze
  positions relative to a presented image. The coordinates have been normalized
  such that a value of 1.0 denotes the furthest onscreen coordinate from the
  central fixation point \textit{towards} the direction of the given image, and
  -1.0 denotes the furthest onscreen coordinate from the central fixation point
  \textit{away from} the direction of the given image. The gray bars in each
  panel mark the boundaries of the presented images. The violin plots are
  broken down by participants' familiarity ratings during the memory phases of
  each experiment, and by each image's relation to the attention cue while that
  image appeared onscreen. Panel \textbf{a.} displays fixation data for the
  Sustained Attention Experiment and Panel \textbf{b.} displays fixation data
  for the Variable Attention Experiment. Figure~\fixation~displays these
  results broken down by participant cohort.}

\label{fig:fixation}
\end{figure}

Given that participants in both experiments rated attended-location images from
both the attended and unattended category as more familiar than
unattended-location images, we wondered whether this was truly a consequence of
covert attention, or whether it might be driven solely by where participants
were looking. In other words, we sought to distinguish the memory effects of
the focus of covert attention from the focus of visual gaze. We used
eyetracking to measure the horizontal positions of the participants' visual
fixations while each pair of composite stimuli appeared onscreen
(Fig.~\ref{fig:fixation}). For 94\% of presentation trials (Sustained Attention
Task: $\frac{4531}{4800}$ trials; Variable Attention Task: $\frac{3972}{4240}$
trials), we collected viable gaze data for analysis. In the remaining trials,
participant movement or other technical issues corrupted the gaze data. We
found that, though they generally followed our instruction to keep their gaze
fixed on the center of the screen, participants in both experiments exhibited a
slight tendency to look towards the side of the screen they were attending to
(Sustained Attention Task: $t(59) = 3.61, p = 6.31 \times 10^{-4}$; Variable
Attention Task: $t(52) = 2.86, p = 6.06 \times 10^{-3}$). When we limited our
analyses to trials with viable gaze data where participants successfully
maintained fixation on the center of the screen within 4.28$^\circ$ visual
angle (i.e., between the rightmost boundary of the left image and the leftmost
boundary of the right image) we observed (numerical) memory benefits of
feature-based attention in both experiments (Sustained Attention Experiment:
$t(59) = 4.73, p = 1.96 \times 10^{-5}, d = 0.67$; Variable Attention
Experiment: $t(52) = 0.64, p = 0.52, d = 0.09$) and location-based attention in
the Sustained Attention Experiment, but not the Variable Attention Experiment
(Sustained Attention Experiment: $t(59) = 0.67, p = 0.51, d = 0.09$; Variable
Attention Experiment: $t(52) = -0.79, p = 0.43, d = 0.11$). These analyses
indicate that although the focus of covert attention can affect where people
are looking, covert attention affects memory encoding beyond what may be
accounted for by gaze position alone.

The preceding results indicate that volitional covert attention affects memory,
and that feature-based and location-based covert attention may come online at
different timescales. We next sought to examine how long the memory effects of
feature-based and location-based covert attention persist. To gain additional
insights into the timecourse of the impact of feature-based and location-based
attention on memory, we used a sliding window analysis to measure how
participants' familiarity ratings of attended, partially attended, unattended,
and novel stimuli varied over the duration of the memory phases of both
experiments. During each block within the memory phases of each experiment,
participants made familiarity judgements about a total of 40 images (20 old
images and 20 novel images). We computed average familiarity ratings for images
appearing in each position of the memory stimulus sequence (positions 1--40),
as a function of how (or whether) they were attended during the preceding
presentation phase (old images) or whether they matched the category of the
most recent attention cue (novel images), for overlapping 20-image sliding
windows (Figs.~\ref{fig:familiarity_timecourse_category},
\ref{fig:familiarity_timecourse_location}, and
\ref{fig:familiarity_timecourse_novel}). We first calculated these values for
each participant, and then we averaged across participants.

\begin{figure*}[tp]
  \centering
  \includegraphics[width=1\textwidth]{figures/Cat_Cued_TimeSeries.pdf}

  \caption{\textbf{Familiarity ratings over time for images that matched the
  cued or uncued image category.} Each curve reflects the average familiarity
  ratings for attended, unattended, and novel images (denoted in the legends on
  the right) within a succession of overlapping 20-image sliding windows. Error
  ribbons denote 95\% confidence intervals, computed across participants.
  Panels \textbf{a.} and \textbf{c.} display results from the Sustained
  Attention Experiment and Panels \textbf{b.} and \textbf{d.} display results
  from the Variable Attention Experiment. The paired horizontal lines at the
  bottom of each panel denote timepoints when the given pair of curves was
  statistically distinguishable (i.e., the topmost line color was statistically
  greater than the bottommost line color at $\alpha = 0.05$, via a paired
  two-tailed $t$-test.) The gray lines in Panels b and d reflect familiarity
  ratings of novel stimuli from both the most recently attended and unattended
  image categories. Figures~\timecourseCuedCategory~and
  \timecourseUncuedCategory~display these results broken down by participant
  cohort.}

\label{fig:familiarity_timecourse_category}
\end{figure*}

>>>>

\newcommand{\gazeLocations}{S1}
\newcommand{\excludedTrials}{S2}
\newcommand{\ratingsByCategory}{S3}

@@

We used the eyetracking data to home in specifically on behavioral effects
related to \textit{covert} attention, as opposed to overt looking effects.
Specifically, we excluded from further analysis any images from trials where
participants shifted their gaze (for any non-zero amount of time) to any part
of the attended composite image during a presentation trial
(see Figs.~\gazeLocations,~and \excludedTrials).

@@

We ran a volitional covert attention experiment with two conditions; in the
sustained attention condition we asked participants to \textit{sustain} the
focus of their attention over a succession of 10 stimulus presentations per
block whereas in the variable attention condition we asked participants to
\textit{vary} their focus of attention with each new stimulus (also for a total
of 10 stimulus presentations per block). Each stimulus comprised a pair of
composite images (one on the left and one on the right side of the display),
where each composite comprised an equal blend of a unique face and a unique
place image. We followed the presentation phases of each experimental block
with a memory phase, where participants performed a recognition memory task by
rating the familiarity of previously experienced and novel face and place
images (see \textit{Experimental paradigm}, Fig.~\ref{fig:paradigm}).

We first wondered whether (and how) shifts in covert attention might affect
participants' ratings during the recognition memory task
(Fig.~\ref{fig:familiarity}). To ensure that our findings were not conflated
with where people were physically looking, we excluded from further analysis
any images presented during trials where the participant's gaze touched on any
part of the attended composite image (see \textit{Eyetraciking},
Figs.~\gazeLocations~and \excludedTrials). For the remaining trials, the
participants kept their gaze focused on a fixation cross at the center of the
screen while \textit{covertly} shifting the focus of their attention to the
cued category component of the composite image on the cued location. In other
words, during these remaining trials, participants' physical (external)
experiences of the face and place components of every presented composite image
remained relatively constant across trials (up to our ability to accurately
measure where participants were looking using the eyetracker).

Simply by encoding their prior experiences into memory, we reasoned that
participants should rate \textit{any} presented images as more familiar than
novel images, regardless of whether they were following the attention cues. We
confirmed that this prediction held in both the sustained ($t(29) = 8.856, p <
0.001$) and variable ($t(22) = 5.144, p < 0.001$) conditions. In addition, to
the extent that participants were following the attention cues, there
\textit{internal} experiences of each image should depend on their internal
focus of attention during each image presentation. For example, we expected
that the attended-category component of the composite image at the attended
location might be better recognized than the other composite image components.
Indeed, participants in both experimental conditions rated these ``fully
attended'' images as more familiar than category-matched image components from
unattended locations (sustained: $t(29) = 6.893, p < 0.001$; variable: $t(22) =
6.938, p < 0.001$), location-matched images from the unattended category
(sustained: $t(29) = 6.710, p < 0.001$; variable: $t(22) = 7.633, p < 0.001$),
unattended images that were neither from the attended category nor the attended
location (sustained: $t(29) = 8.470, p < 0.001$; variable: $t(22) = 7.256, p <
0.001$), or novel images they had never seen before (sustained: $t(29) =
10.259, p < 0.001$; variable: $t(22) = 7.874, p < 0.001$). We also observed a
``boost'' in familiarity over novel images for attended \textit{or} unattended
images from the cued category (sustained: $t(29) = 10.578, p < 0.001$;
variable: $t(22) = 7.354, p < 0.001$) and location (sustained: $t(29) = 8.917,
p < 0.001$; variable: $t(22) = 5.726, p < 0.001$).

We also wondered whether the ways participants attended to or remembered the
images might depend on image-specific properties like the images' categories.
We repeated the analyis displayed in Figure~\ref{fig:familiarity} separately
for face and place images (Fig.~\ratingsByCategory). The same general patterns
held for each category, as when we combined across the two stimulus categories
as reported above. For example, fully attended face and place images were both
rated as more familiar than the category-matched images from the unattended
location (sustained: $t$s$(29) \geq 3.36, p$s$ \leq 0.002$; variable: $t$s$(22)
\geq 6.205, p$s$ \leq 0.001$), attended-location images from the unattended
category (sustained: $t$s$(29) \geq 5.886, p$s$ \leq 0.001$; variable:
$t$s$(29) \geq 4.277, p$s$ \leq 0.001$), images from the unattended category
and location (sustained: $t$s$(29) \geq 6.628, p$s$ \leq 0.001$; variable:
$t$s$(29) \geq 5.624, p$s$ \leq 0.001$), and novel images (sustained: $t$s$(29)
\geq 5.987, p$s$ \leq 0.001$; variable: $t$s$(29) \geq 5.132, p$s$ \leq
0.001$). Taken together, the above results suggest that what we remember is
guided in part by what we attended to, even after accounting for where we look
or what specifically we are looking at.

Splitting participants' responses to face versus place images also revealted
that participants often rated attended (and partially attended) place images as
more familiar than attention-matched face images (compare dark versus light
bars in Fig.~\ratingsByCategory). We hypothesized that this might be
explainable by some property of the relevant cognitive processes or by
properties of the stimuli themselves. To help elucidate this distinction, we
examined individual exemplars of the face and place images used in our paradigm
(Fig.~\ref{fig:stimuli}A). By design, the face images had consistent head
sizes, viewing angles, expressions, and so on. In contrast, the place images
varied more substantially across images. For example, some place images
depicted human-made structures; others depicted natural scenes; some depicted
indoor views; others depicted outdoor views; etc. This can also be seen by
averaging the pixel intensity values across images, separately for the face and
place stimuli (Fig.~\ref{fig:stimuli}B). Whereas the average face image retains
many of the landmarks characteristic of most faces (e.g., clearly defined hair,
eyes, nose, mouth, head shape, etc.), the average place image does not show
place-specific features as clearly, aside from a general tendancy for the tops
of place images to be lighter than the bottoms of place images. We also
computed the pairwise similarities across images from each stimulus category
(Fig.~\ref{fig:stimuli}C) and found that face images tended to be much more
similar to each other than place images (Fig.~\ref{fig:stimuli}D; $t(115258) =
254.764, p < 0.001$). This analysis indicated to us that our experimental
paradigm was not well-suited to identifying cognitively meaningful stimulus
category differences, since participants' category-specific judgements may be
confounded with within-category image similarity differences.

\begin{figure*}[tp]
	\centering
	\includegraphics[width=\textwidth]{figs/stimuli}

  \caption{\textbf{Stimulus examples and properties.} \textbf{A. Example images
  from each stimulus category.} Randomly chosen subsets of 20 face images
  (left) and 20 place images (right) are displayed. \textbf{B. Across-image
  averages.} Each panel displays the average image, taken across all 360 face
  images (left) and 360 place images (right). \textbf{C. Pairwise
  correlations.} Each row and column of the matrices displays the correlation
  (across pixels) in intensity values for one pair of face images (left) or
  place images (right). \textbf{D. Average pairwise correlations.} The bar
  heights denote the average pairwise correlations between face and place
  images. Error bars denote across-pair bootstrap-estimated 95\%
  confidence intervals.}

  \label{fig:stimuli}
  \end{figure*}

Next, we turned to identifying potential differences in participants' behaviors
across the two experimental conditions. The main difference between the
conditions' procedures was in how long participants were asked to maintain the
same focus of category-based and location-based attention, across successive
image presentations. Therefore, differences in participants' behaviors across
the two conditions might reflect differences in the timescales of the relevant
cognitive processes. We compared participants' familiarity ratings for images
at each attention level across the two conditions. We saw no evidence that
people rated fully attended images ($t(51) = -0.649, p = 0.519$),
attended-category (but not location) images ($t(51) = 1.163, p = 0.250$),
unattended category and location images ($t(51) = -1.600, p = 0.116$), or novel
images ($t(51) = -0.435, p = 0.665$) differently across the two conditions.
However, participants in the variable attention condition rated
attended-location (but unattended category) images as more familiar than
participants in the sustained attention condition ($t(51) = 2.174, p = 0.034$).
This suggests that the timescales over which location-based versus
category-based attention tend to operate may differ.

@@

% We considered each image we probed in the memory phase with respect to whether
% (and how) the participant had experienced and/or attended to that image during
% the presentation phase (Fig.~\ref{fig:familiarity}). For example, if a given
% face image had been presented as part of the composite image on the right of
% the screen following an \texttt{attend to the face on the right} instruction,
% then that face image would have been the focus of the participant's attention
% during one stimulus presentation. The face image on the left (unattended side)
% during that trial would have been captured by feature-based attention, but not
% by location-based attention; and the place image (unattended category) on the
% right during that trial would have been captured by location-based attention,
% but not by feature-based attention. The place image on the left during that
% trial remained outside of the focus of both feature-based and location-based
% attention. In this way, we categorized each of the images that participants
% experienced during the presentation phases of each experiment in terms of
% whether they fell under the scope of feature-based and/or location-based
% attention. We also categorized novel face and place images that we asked
% participants to judge during the memory phase (i.e., images that the
% participant hadn't seen before) as belonging to the attention-cued or uncued
% category. These novel images were intended to serve as a baseline for
% comparison. For example, we sought to measure how people rated the familiarity
% of new images in general. We also used differences in familiarity ratings
% between novel images of the attention-cued versus uncued categories to evaluate
% potential \textit{prospective} affects of modulating the focus of feature-based
% attention.

% \begin{figure*}[tp]
%   \centering
%   \includegraphics[width=0.75\textwidth]{figs/familiarity_by_attention_level}

%   \caption{\textbf{Familiarity ratings for attended, unattended, and novel
%   stimuli.} The bars display the average familiarity ratings participants gave
%   to images from the same category and location as the attention cue (full
%   attention), the same category (but opposite location) as the attention cue
%   (attended category), the same location (but opposite category) as the
%   attention cue (attended location), the opposite category and location as the
%   attention cue (unattended), or novel images. The left panel displays
%   familiarity ratings from the sustained attention condition and the right
%   panel displays familiarity ratings from the variable attention condition. All
%   error bars denote across-participant bootstrap-estimated 95\% confidence
%   intervals. See Figure~\ratingsByCategory~for analogous results broken down by
%   stimulus category.}

% \label{fig:familiarity}
% \end{figure*}

% Participants in both experiments rated stimuli they attended as more familiar
% than unattended or novel stimuli (Fig.~\ref{fig:familiarity}; Sustained
% Attention Experiment: $t(59) = 13.42, p = 1.39 \times 10^{-19}, d = 1.73$;
% Variable Attention Experiment: $t(52) = 12.40, p = 3.70 \times 10^{-17}, d =
% 1.70$). Participants in both experiments also rated attended scene stimuli as
% more familiar than attended face stimuli (Sustained Attention Experiment:
% $t(59) = 7.69, p = 1.85 \times 10^{-10}, d = 0.99$; Variable Attention
% Experiment: $t(52) = 7.08, p = 3.63 \times 10^{-9}, d = 0.97$). These findings
% indicate that aspects of experience that are captured by the focus of attention
% are easier to recognize later, and that attention may preferentially benefit
% memory for some aspects of experience to a greater extent than other aspects of
% experience.

% Participants in the Sustained Attention Experiment experiment also rated
% unattended images that matched the attended category as more familiar than
% unattended images that did not match the attended category (\textit{incidental
% feature-based attention}; $t(59) = 7.17, p = 1.40 \times 10^{-9}, d = 0.93$).
% This pattern did not hold for participants in the Variable Attention Experiment
% ($t(52) = 0.44, p = 0.66, d = 0.06$). Participants in both experiments
% displayed an incidental effect of location-based attention, whereby they rated
% unattended-category images at the attended location as more familiar than
% unattended-category images at the unattended location (Sustained Attention
% Experiment: $t(59) = 4.32, p = 6.05 \times 10^{-5}, d = 0.56$; Variable
% Attention Experiment: $t(52) = 2.57, p = 0.01, d = 0.35$). This indicates that
% aspects of our experience that incidentally overlap with the focus of our
% location-based attention are easier to recognize later, even if those aspects
% do not fall within the intentional focus of our feature-based attention.

% We defined the \textit{memory benefit of feature-based attention} as the
% difference in mean familiarity ratings of attended stimuli and unattended
% stimuli from the attended category versus the mean familiarity ratings of
% unattended stimuli from the unattended category. Similarly, we defined the
% \textit{memory benefit of location-based attention} as the difference in mean
% familiarity ratings of attended stimuli and unattended stimuli from the
% attended location versus the mean familiarity ratings of unattended stimuli
% from the unattended location. Participants in the Sustained Attention
% Experiment displayed a larger memory benefit of feature-based attention than
% location-based attention ($t(59) = 3.25, p = 1.91 \times 10^{-3}, d = 0.42$),
% whereas participants in the Variable Attention Experiment displayed a larger
% memory benefit of location-based attention ($t(52) = 1.70, p = 9.99 \times
% 10^{-2}, d = 0.23$). This appeared to be driven in part by a response bias
% towards attended-category images. For example, relative to the most recent
% attention cue, participants in the Sustained Attention Experiment rated
% attended-category novel images as more familiar than unattended-category novel
% images ($t(59) = 7.04, p = 2.35 \times 10^{-9}, d = 0.91$), whereas
% participants in the Variable Attention Experiment responded similarly to
% attended-category and unattended-category novel images ($t(52) = 0.23, p =
% 0.82, d = 0.03$). The size of this attended-category response bias for novel
% images was reliably smaller than the memory benefits of feature-based attention
% in both experiments (Sustained Attention Experiment: $t(59) = 7.59, p = 2.68
% \times 10^{-10}, d = 0.98$; Variable Attention Experiment: $t(52) = 6.74, p =
% 1.31 \times 10^{-8}, d = 0.93$). This indicates that tuning the focus of
% feature-based attention with respect to how ongoing experience is encoded in
% memory takes longer than the duration of a single stimulus presentation
% (including fixation, roughly 5.5~s). Further, once feature-based attention
% comes ``online,'' it affects how \textit{new} stimuli are processed.

% We also compared the relative sizes of the memory benefits of feature-based
% attention and location-based attention across the two experiments. Participants
% in the Sustained Attention Experiment displayed a larger difference between
% these two effects than did participants in the Variable Attention Experiment
% ($t(111) = 3.48, p = 7.24 \times 10^{-4}, d = 0.66$). Further, the difference
% in mean familiarity ratings of attended-location images versus unattended
% images was greater for participants in the Variable Attention Experiment
% ($t(111) = 2.32, p = 0.02, d = 0.44$). The memory benefit of location-based
% attention reliably distinguished between attended-category versus unattended
% category images for participants in both experiments (Sustained Attention
% Experiment: $t(59) = 9.50, p = 1.68 \times 10^{-13}, d = 1.23$; Variable
% Attention Experiment: $t(52) = 9.72, p = 2.76 \times 10^{-13}, d = 1.33$). This
% indicates that tuning the focus of location-based attention with respect to how
% ongoing experience is encoded into memory occurs faster than the duration of a
% single stimulus presentation.

% We also compared reaction times to probes from the attended versus unattended
% side. Participants in both experiments responded slightly faster on average to
% attended-side probes, although the differences did not reach significance
% (Sustained Attention Experiment: $t(59) = 1.03, p = 0.31, d = 0.13$; Variable
% Attention Experiment: $t(52) = 0.65, p = 0.52, d = 0.09$). Participants in the
% Sustained Attention Experiment displayed a slightly larger (and not
% statistically significant) difference in attended-side versus unattended-side
% reaction times than participants in the Variable Attention Experiment ($t(111)
% = .87, p = 0.38, d = 0.20$). Because none of these reaction time effects were
% statistically reliable, we are unable to draw meaningful conclusions about
% reaction times to the attention probes on the attended versus unattended side,
% nor are we able to draw meaningful conclusions about differences in those
% reaction times that might be due to sustained versus variable attention.

% \begin{figure}[tp]
%   \centering
%   \includegraphics[width=\textwidth]{figures/exps1_2_xgaze.pdf}

%   \caption{\textbf{Horizontal coordinates of visual fixations.} Each violin
%   plot displays a distribution of participant-wise average horizontal gaze
%   positions relative to a presented image. The coordinates have been normalized
%   such that a value of 1.0 denotes the furthest onscreen coordinate from the
%   central fixation point \textit{towards} the direction of the given image, and
%   -1.0 denotes the furthest onscreen coordinate from the central fixation point
%   \textit{away from} the direction of the given image. The gray bars in each
%   panel mark the boundaries of the presented images. The violin plots are
%   broken down by participants' familiarity ratings during the memory phases of
%   each experiment, and by each image's relation to the attention cue while that
%   image appeared onscreen. Panel \textbf{a.} displays fixation data for the
%   Sustained Attention Experiment and Panel \textbf{b.} displays fixation data
%   for the Variable Attention Experiment. Figure~\fixation~displays these
%   results broken down by participant cohort.}

% \label{fig:fixation}
% \end{figure}

% Given that participants in both experiments rated attended-location images from
% both the attended and unattended category as more familiar than
% unattended-location images, we wondered whether this was truly a consequence of
% covert attention, or whether it might be driven solely by where participants
% were looking. In other words, we sought to distinguish the memory effects of
% the focus of covert attention from the focus of visual gaze. We used
% eyetracking to measure the horizontal positions of the participants' visual
% fixations while each pair of composite stimuli appeared onscreen
% (Fig.~\ref{fig:fixation}). For 94\% of presentation trials (Sustained Attention
% Task: $\frac{4531}{4800}$ trials; Variable Attention Task: $\frac{3972}{4240}$
% trials), we collected viable gaze data for analysis. In the remaining trials,
% participant movement or other technical issues corrupted the gaze data. We
% found that, though they generally followed our instruction to keep their gaze
% fixed on the center of the screen, participants in both experiments exhibited a
% slight tendency to look towards the side of the screen they were attending to
% (Sustained Attention Task: $t(59) = 3.61, p = 6.31 \times 10^{-4}$; Variable
% Attention Task: $t(52) = 2.86, p = 6.06 \times 10^{-3}$). When we limited our
% analyses to trials with viable gaze data where participants successfully
% maintained fixation on the center of the screen within 4.28$^\circ$ visual
% angle (i.e., between the rightmost boundary of the left image and the leftmost
% boundary of the right image) we observed (numerical) memory benefits of
% feature-based attention in both experiments (Sustained Attention Experiment:
% $t(59) = 4.73, p = 1.96 \times 10^{-5}, d = 0.67$; Variable Attention
% Experiment: $t(52) = 0.64, p = 0.52, d = 0.09$) and location-based attention in
% the Sustained Attention Experiment, but not the Variable Attention Experiment
% (Sustained Attention Experiment: $t(59) = 0.67, p = 0.51, d = 0.09$; Variable
% Attention Experiment: $t(52) = -0.79, p = 0.43, d = 0.11$). These analyses
% indicate that although the focus of covert attention can affect where people
% are looking, covert attention affects memory encoding beyond what may be
% accounted for by gaze position alone.

% The preceding results indicate that volitional covert attention affects memory,
% and that feature-based and location-based covert attention may come online at
% different timescales. We next sought to examine how long the memory effects of
% feature-based and location-based covert attention persist. To gain additional
% insights into the timecourse of the impact of feature-based and location-based
% attention on memory, we used a sliding window analysis to measure how
% participants' familiarity ratings of attended, partially attended, unattended,
% and novel stimuli varied over the duration of the memory phases of both
% experiments. During each block within the memory phases of each experiment,
% participants made familiarity judgements about a total of 40 images (20 old
% images and 20 novel images). We computed average familiarity ratings for images
% appearing in each position of the memory stimulus sequence (positions 1--40),
% as a function of how (or whether) they were attended during the preceding
% presentation phase (old images) or whether they matched the category of the
% most recent attention cue (novel images), for overlapping 20-image sliding
% windows (Figs.~\ref{fig:familiarity_timecourse_category},
% \ref{fig:familiarity_timecourse_location}, and
% \ref{fig:familiarity_timecourse_novel}). We first calculated these values for
% each participant, and then we averaged across participants.

% \begin{figure*}[tp]
%   \centering
%   \includegraphics[width=1\textwidth]{figures/Cat_Cued_TimeSeries.pdf}

%   \caption{\textbf{Familiarity ratings over time for images that matched the
%   cued or uncued image category.} Each curve reflects the average familiarity
%   ratings for attended, unattended, and novel images (denoted in the legends on
%   the right) within a succession of overlapping 20-image sliding windows. Error
%   ribbons denote 95\% confidence intervals, computed across participants.
%   Panels \textbf{a.} and \textbf{c.} display results from the Sustained
%   Attention Experiment and Panels \textbf{b.} and \textbf{d.} display results
%   from the Variable Attention Experiment. The paired horizontal lines at the
%   bottom of each panel denote timepoints when the given pair of curves was
%   statistically distinguishable (i.e., the topmost line color was statistically
%   greater than the bottommost line color at $\alpha = 0.05$, via a paired
%   two-tailed $t$-test.) The gray lines in Panels b and d reflect familiarity
%   ratings of novel stimuli from both the most recently attended and unattended
%   image categories. Figures~\timecourseCuedCategory~and
%   \timecourseUncuedCategory~display these results broken down by participant
%   cohort.}

% \label{fig:familiarity_timecourse_category}
% \end{figure*}

% Throughout the memory phase, participants from both experiments rated attended
% images as more familiar than category-matched unattended and novel images
% (Fig.~\ref{fig:familiarity_timecourse_category}, top row). This indicates that
% the memory benefits of attending to specific stimulus features (versus
% category-general features) persist for at least as long as the duration of the
% memory phase (2~min). Early in the memory phase, participants in the Variable
% Attention Experiment also rated category-matched unattended images as more
% familiar than novel images. This pattern did not hold for participants in the
% Sustained Attention Experiment. The short-lived boost in familiarity for
% attended-category images at the unattended location following variable (but not
% sustained) attention suggests that the timecourse of feature-based attention's
% impact on memory depends on how long the focus of feature-based attention has
% been held.
